\documentclass{ximera}

\title{Function Review Activity}
\author{MATH 425: Calculus I}

\begin{document}
\begin{abstract}
    Working with the peers in your group, solve the following problems. Make sure to show and justify all your work. Make sure everyone in the group understands the solution and participates. Be prepared to report your answers to the whole class. 
\end{abstract}
\maketitle


\begin{exercise}

 In this problem, let $x$ denote the length of the base of a rectangle. Suppose the height of the rectangle is always half the length of its base.
    \begin{enumerate}
        \item Draw a labeled sketch of the rectangle.
        \item Let $A(x)$ be a function that gives the area of the rectangle whose base has length $x$. Find a formula for $A(x)$ as a function of $x$.
        \item What is the domain of the function $A(x)$? Explain your answer.
        \item Let $P(x)$ be a function that gives the perimeter of the rectangle whose base has length $x$. Find a formula for $P(x)$ as a function of $x$.
       % Using the work you did on your homework, 
       \item Write the area of the rectangle, $A(x)$, in terms of the perimeter $P(x)$. 
    \end{enumerate}
   \end{exercise}

\begin{exercise}
   Determine the domain of the function $g(x)=\frac{\sqrt{x-1}}{x^2-9}$.
\end{exercise}

\begin{exercise}
     Suppose \[ f(x)=\begin{cases} 
     3x-2 & x< 0 \\
      x^2+1 & 0< x\leq 4 \\
      0 &  4< x 
   \end{cases}
\]
   \begin{enumerate}
    \item Evaluate $f(-2)$, $f(0)$, $f(3)$ and $f(5)$.
    \item Sketch a graph of the function. 
   \end{enumerate}
 
\end{exercise}

\begin{exercise}
   Consider the following functions.  Classify the functions in the following ways, and be prepared to defend your answers. \\
    $f(x)=4^x$\\
    $g(x)=x^6$\\
    $h(x)=\frac{1+2x}{1+\sqrt{x}}$\\
    $F(t)=1-2.1t+3t^2$\\
    $G(t)=\log_2(t)$\\
    $H(t)=\sqrt[3]{t}$\\
    $P(q)=\frac{2q^3}{1-q^3}$\\
    $R(m)=\sin(m)$\\
    $C(z)=z$
  \begin{enumerate}
      \item Are any of these functions classified as linear functions?  Which one(s)?
      \item Are any of these functions classified as polynomial functions?  Which one(s)?
      \item Are any of these functions classified as power functions functions?  Which one(s)?
      \item Are any of these functions classified as rational functions functions?  Which one(s)?
      \item Are any of these functions classified as algebraic functions?  Which one(s)?
  \end{enumerate}
\end{exercise}

\end{document}